---
title:  |
        | RadPredict R model code
        | Jared Tavares
date: "`r Sys.Date()`"
output: 
  bookdown::pdf_document2:
    toc: false
    number_sections: false
---

```{r setup, include=F}

# Clear the environment before you begin
rm(list=ls())

# setwd("C:/Users/nicol/OneDrive/Radiological predictors of PCP/Data analysis/R")

library("knitr")
library(kableExtra)
knitr::opts_chunk$set(echo = T, warning = F, message = F, fig.pos = 'H')
options(digits=4)
require("latex2exp")
library(readxl)
library(dplyr)
library(readr)
library(tidyr)
library(ggplot2)
library(earth)
library(MASS)
library(glmnet)
library(pROC)
library(xtable)
library(rrpack)
library(Metrics)
library(stringr)
library(tidyverse)
library(car)
library(broom)
library(Hmisc)
library(caret)
require(rms)
library(purrr)
library(epiR)

```


***

# Loading and setting up the data


The code for setting up each of the individual custom variables was in the previous RadPredict RMD file. Please refer to that file if you would like to see how to create the variables. The code below is for generating the values for the tables for both continuous and categorical variables. The values for all columns are provided by the output. 

```{r}

#write.csv(data, "data_clean.csv", row.names = FALSE)

```


```{r}

set.seed(123)

# Load the data - this is from RedCap
data_clean = read.csv("data_clean.csv", header = T)

# Exclude the patients that need to be excluded
data_clean = data_clean %>% filter(is.na(excl_1))

data_clean$response = ifelse(data_clean$micro_result.factor == "Positive", 1, 0) # 1 for positive, 0 for negative (updated)

```



```{r}
##
## Variables for the venn diagram are below
##

# Define the number lists for each condition
tb_numbers <- c(1, 23)
tb_numbers_lab = c(1)

bp_numbers <- c(2, 22)
vp_numbers <- c(4, 5)
pe_numbers <- c(8, 9)
other_numbers <- c(16, 6, 17)


data_clean$empiricdx_1[data_clean$empiricdx_1 %in% tb_numbers]

cbind(data_clean$empiricdx_1, data_clean$empiricdx_1.factor)


# Define the condition variables
data_clean$tb <- ifelse(!is.na(data_clean$empiricdx_1) & data_clean$empiricdx_1 %in% tb_numbers, 1,
                        ifelse(!is.na(data_clean$empiricdx_2) & data_clean$empiricdx_2 %in% tb_numbers, 1,
                               ifelse(!is.na(data_clean$empiricdx_3) & data_clean$empiricdx_3 %in% tb_numbers, 1, 
                                      ifelse(!is.na(data_clean$labconfirmdx_1) & data_clean$labconfirmdx_1 %in% tb_numbers_lab ,1,0))))

data_clean$bacterialPneumonia <- ifelse(!is.na(data_clean$empiricdx_1) & data_clean$empiricdx_1 %in% bp_numbers, 1,
                                        ifelse(!is.na(data_clean$empiricdx_2) & data_clean$empiricdx_2 %in% bp_numbers, 1,
                                               ifelse(!is.na(data_clean$empiricdx_3) & data_clean$empiricdx_3 %in% bp_numbers, 1, 0)))

data_clean$viralPneumonia <- ifelse(!is.na(data_clean$empiricdx_1) & data_clean$empiricdx_1 %in% vp_numbers, 1,
                                    ifelse(!is.na(data_clean$empiricdx_2) & data_clean$empiricdx_2 %in% vp_numbers, 1,
                                           ifelse(!is.na(data_clean$empiricdx_3) & data_clean$empiricdx_3 %in% vp_numbers, 1, 0)))

data_clean$pulmonaryOedema <- ifelse(!is.na(data_clean$empiricdx_1) & data_clean$empiricdx_1 %in% pe_numbers, 1,
                                     ifelse(!is.na(data_clean$empiricdx_2) & data_clean$empiricdx_2 %in% pe_numbers, 1,
                                            ifelse(!is.na(data_clean$empiricdx_3) & data_clean$empiricdx_3 %in% pe_numbers, 1, 0)))

data_clean$other <- ifelse(!is.na(data_clean$empiricdx_1) & data_clean$empiricdx_1 %in% other_numbers, 1,
                           ifelse(!is.na(data_clean$empiricdx_2) & data_clean$empiricdx_2 %in% other_numbers, 1,
                                  ifelse(!is.na(data_clean$empiricdx_3) & data_clean$empiricdx_3 %in% other_numbers, 1, 0)))


##
## Custom variables included in the first tables (check these before using them for the tables below)
## Any other custom variables can be created in a similar way.
##


data_clean$age = as.integer(as.numeric(difftime(as.Date(data_clean$date_admission), 
                                                as.Date(data_clean$dob), units = "days"))/365.25)

data_clean$sats_ra[which(data_clean$sats_ra == 0)] = NA
data_clean$pa02_ra[which(data_clean$pa02_ra == 0)] = NA
hypox_1 = which(data_clean$sats_ra < 90)
hypox_2 = which(data_clean$pa02_ra < 7.8)
hypox_3 = which(data_clean$comb_pf <= 300) # Changed this

hypox = unique(c(hypox_1, hypox_2, hypox_3))
hypox[which(is.na(data_clean$comb_pf))] = NA

hypox

data_clean$hypox = ifelse(1:104 %in% hypox, "Yes", "No")

describe(data_clean$hypox)


# Check that they went to icu after doing this
dur_admission_1 <- as.integer(as.numeric(difftime(as.Date(data_clean$date_discharge), as.Date(data_clean$date_admission) , units = "days")))
dur_admission_2 <- as.integer(as.numeric(difftime(as.Date(data_clean$date_inhop_death), as.Date(data_clean$date_admission), units = "days")))

dur_admission_1[is.na(dur_admission_1)] = dur_admission_2[is.na(dur_admission_1)]

# Duration of admission
data_clean$dur_admission = dur_admission_1


# Time to initiation of PCP treatment (days) - Check some of them
time_init_pcp <- as.integer(as.numeric(difftime(as.Date(data_clean$date_pcp_rx), as.Date(data_clean$date_admission) , units = "days")))
data_clean$time_init_pcp = time_init_pcp


# Number of days from admission to ICU admission
days_admit_to_icu <- as.integer(as.numeric(difftime(as.Date(data_clean$date_icu_admit), as.Date(data_clean$date_admission) , units = "days")))
data_clean$days_admit_to_icu = days_admit_to_icu

# Define mapping
mapping = list(
  "Tertiary" = "GSH",
  "District/Regional" = c("VHW", "MPH", "NSH")
)

# Update `data_clean$hosp` based on mapping
data_clean$hosp_cat <- sapply(data_clean$hosp, function(hosp_string) {
  if (str_detect(hosp_string, mapping[["Tertiary"]])) {
    return("Tertiary")
  } else if (any(str_detect(hosp_string, mapping[["District/Regional"]]))) {
    return("District/Regional")
  } else {
    return(NA)  # return NA if no match is found
  }
})

# Co-diagnosis

# Specify the list of numbers to check for
check_numbers <- c(1, 2, 3, 4, 5, 6, 8, 9, 16, 17, 23)

# Create the new variable
data_clean$codiag <- ifelse(!is.na(data_clean$empiricdx_1) & data_clean$empiricdx_1 %in% check_numbers, 1,
                            ifelse(!is.na(data_clean$empiricdx_2) & data_clean$empiricdx_2 %in% check_numbers, 1,
                                   ifelse(!is.na(data_clean$empiricdx_3) & data_clean$empiricdx_3 %in% check_numbers, 1, 0)))

data_clean$codiag = as.factor(data_clean$codiag) # 0 for no, 1 for yes

data_clean$codiag

# Detectable and undetectable VL
describe(data_clean$vl)
data_clean$vl = str_trim(data_clean$vl)
data_clean$undetectable_vl = ifelse(data_clean$vl == "LDL" | data_clean$vl == "< 20", "Yes", "No")

data_clean$undetectable_vl

data_clean[, c("vl", "undetectable_vl")]


data_clean$detectable_vl = as.numeric(data_clean$vl)
data_clean$detectable_vl

# Check for empty strings in the pf column and replace them with the corresponding value from imputedpf
data_clean$comb_pf <- ifelse(data_clean$pf == "", data_clean$imputedpf, data_clean$pf)
data_clean$comb_pf = as.numeric(data_clean$comb_pf)
data_clean$comb_pf

# Create the severity score
data_clean$severity.score = 0

gg_consol.score = data_clean$parenchy_yn___2 * 3 + data_clean$parenchy_yn___3 * 3
gg_consol.score = ifelse(gg_consol.score == 6, 3, gg_consol.score)

parenchymal_df = as.data.frame(cbind(data_clean$parenchy_yn___7 * 1, # Normal
                                     data_clean$parenchy_yn___1 * 2, # R or RT
                                     gg_consol.score # ground glass or consol
))
# Can't get more than 3 points - so ground glass or consol then 3 pts

parenchymal_df <- parenchymal_df %>%
  rowwise() %>%
  mutate(sum = sum(c_across(everything())))

parenchymal_df

# Points for involvement for each zone - set 0 for NA
data_clean$tot_zones[is.na(data_clean$tot_zones)] = 0

data_clean$tot_zones

# Point for diffuse descriptor? (check this)
diffuse_desc = ifelse((data_clean$retic_patt_ext.factor == "Diffuse") |
                        (data_clean$gg_ext.factor == "Diffuse") |
                        (data_clean$consol_ext.factor == "Diffuse"), 1, 0)

diffuse_desc = ifelse(is.na(diffuse_desc), 0, diffuse_desc) # Get rid of the NA values

data_clean$severity.score = parenchymal_df$sum + data_clean$tot_zones + diffuse_desc

#CD4 category variable
data_clean$CD4_lt50 = ifelse (data_clean$cd4 < 50,"yes","no")

# PF ratio
data_clean$pf
data_clean$imputedpf

# Check for empty strings in the pf column and replace them with the corresponding value from imputedpf
data_clean$comb_pf <- ifelse(data_clean$pf == "", data_clean$imputedpf, data_clean$pf)
data_clean$comb_pf = as.numeric(data_clean$comb_pf)
data_clean$comb_pf

#alb_new
data_clean$alb_new <- ifelse(data_clean$alb != "not done", data_clean$alb, NA)
data_clean$alb_new = as.numeric(data_clean$alb_new)
data_clean$alb_new

# Any calcification - variables were slightly different in latest RedCap download
data_clean$any_calc = ifelse((data_clean$retic_calcs.factor == "Yes") |
                               (data_clean$gg_calcs.factor == "Yes") |
                               (data_clean$consol_calcs.factor == "Yes") |
                               (data_clean$cysts_calcs.factor == "Yes") |
                               (data_clean$cav_calcs.factor == "Yes") |
                               (data_clean$other_op_calcs.factor == "Yes"), "Yes", "No")
```


```{r}
# Variables for table 4

# Any reticular pattern - not sure whether or not to include all parenchy_yn variable or just the first one.
data_clean$parenchy_yn___1.factor 

# Nodules present - lots of NA values
data_clean$nodules.factor

# Nodules charac
data_clean$nodules_charac.factor

# Retic pattern ext
data_clean$retic_patt_ext.factor

# Retic pattern zones
data_clean$retic_patt_zones___1.factor # RUZ
data_clean$retic_patt_zones___2.factor # RMZ
data_clean$retic_patt_zones___3.factor # RLZ
data_clean$retic_patt_zones___4.factor # LUZ
data_clean$retic_patt_zones___5.factor # LMZ
data_clean$retic_patt_zones___6.factor # LLZ

data_clean$zone_count = as.integer(data_clean$retic_patt_zones___1.factor == "Checked") +
                         as.integer(data_clean$retic_patt_zones___2.factor == "Checked") +
                         as.integer(data_clean$retic_patt_zones___3.factor == "Checked") +
                         as.integer(data_clean$retic_patt_zones___4.factor == "Checked") +
                         as.integer(data_clean$retic_patt_zones___5.factor == "Checked") +
                         as.integer(data_clean$retic_patt_zones___6.factor == "Checked")

data_clean$zone_count

data_clean$zone_count[which(data_clean$zone_count == 0)] = NA  

data_clean$zone_category = cut(data_clean$zone_count, breaks = c(0, 1.5, 4.5, 6), 
                                labels = c("< 2", "2-4", "> 4"), include.lowest = TRUE, right = TRUE)

data_clean$zone_category

# Proportions of records with any involvement in each of the three general zones (as a proportion of all participants):

# Upper zone (RUZ or LUZ)
upper_zone_proportion_r = mean(data_clean$retic_patt_zones___1.factor == "Checked" | 
                              data_clean$retic_patt_zones___4.factor == "Checked")

# Mid zone (RMZ or LMZ)
mid_zone_proportion_r = mean(data_clean$retic_patt_zones___2.factor == "Checked" | 
                            data_clean$retic_patt_zones___5.factor == "Checked")

# Lower zone (RLZ or LLZ)
lower_zone_proportion_r = mean(data_clean$retic_patt_zones___3.factor == "Checked" | 
                              data_clean$retic_patt_zones___6.factor == "Checked")

# Upper zone
data_clean$upper_zone_proportion_r = data_clean$retic_patt_zones___1.factor == "Checked" | 
                              data_clean$retic_patt_zones___4.factor == "Checked"

# Mid zone (RMZ or LMZ)
data_clean$mid_zone_proportion_r = data_clean$retic_patt_zones___2.factor == "Checked" | 
                            data_clean$retic_patt_zones___5.factor == "Checked"

# Lower zone (RLZ or LLZ)
data_clean$lower_zone_proportion_r = data_clean$retic_patt_zones___3.factor == "Checked" | 
                              data_clean$retic_patt_zones___6.factor == "Checked"

data_clean$retic_patt_zones___3.factor == "Checked" | 
                              data_clean$retic_patt_zones___6.factor == "Checked"

# Print the proportions
cat("Proportion of records with Upper zone involvement:", upper_zone_proportion, "\n")
cat("Proportion of records with Mid zone involvement:", mid_zone_proportion, "\n")
cat("Proportion of records with Lower zone involvement:", lower_zone_proportion, "\n")

# A histogram to visualise this better
histogram(data_clean$zone_category)

### GG Zones ###

data_clean$gg_zones___1.factor # RUZ
data_clean$gg_zones___2.factor # RMZ
data_clean$gg_zones___3.factor # RLZ
data_clean$gg_zones___4.factor # LUZ
data_clean$gg_zones___5.factor # LMZ
data_clean$gg_zones___6.factor # LLZ

data_clean$zone_count_gg = as.integer(data_clean$gg_zones___1.factor == "Checked") +
                         as.integer(data_clean$gg_zones___2.factor == "Checked") +
                         as.integer(data_clean$gg_zones___3.factor == "Checked") +
                         as.integer(data_clean$gg_zones___4.factor == "Checked") +
                         as.integer(data_clean$gg_zones___5.factor == "Checked") +
                         as.integer(data_clean$gg_zones___6.factor == "Checked")

data_clean$zone_count_gg

data_clean$zone_category_gg = cut(data_clean$zone_count_gg, breaks = c(0, 1.5, 4.5, 6), 
                                labels = c("< 2", "2-4", "> 4"), include.lowest = TRUE, right = TRUE)


# A histogram to visualise this better
histogram(data_clean$zone_category_gg)

# Proportions of records with any involvement in each of the three general zones (as a proportion of all participants):

# Upper zone (RUZ or LUZ)
upper_zone_proportion = mean(data_clean$gg_zones___1.factor == "Checked" | 
                              data_clean$gg_zones___4.factor == "Checked")

# Mid zone (RMZ or LMZ)
mid_zone_proportion = mean(data_clean$gg_zones___2.factor == "Checked" | 
                            data_clean$gg_zones___5.factor == "Checked")

# Lower zone (RLZ or LLZ)
lower_zone_proportion = mean(data_clean$gg_zones___3.factor == "Checked" | 
                              data_clean$gg_zones___6.factor == "Checked")

# Print the proportions
cat("Proportion of records with Upper zone involvement:", upper_zone_proportion, "\n")
cat("Proportion of records with Mid zone involvement:", mid_zone_proportion, "\n")
cat("Proportion of records with Lower zone involvement:", lower_zone_proportion, "\n")


### Consolidation ###

data_clean$consol_zones___1.factor # RUZ
data_clean$consol_zones___2.factor # RMZ
data_clean$consol_zones___3.factor # RLZ
data_clean$consol_zones___4.factor # LUZ
data_clean$consol_zones___5.factor # LMZ
data_clean$consol_zones___6.factor # LLZ

data_clean$zone_count_c = as.integer(data_clean$consol_zones___1.factor == "Checked") +
                         as.integer(data_clean$consol_zones___2.factor == "Checked") +
                         as.integer(data_clean$consol_zones___3.factor == "Checked") +
                         as.integer(data_clean$consol_zones___4.factor == "Checked") +
                         as.integer(data_clean$consol_zones___5.factor == "Checked") +
                         as.integer(data_clean$consol_zones___6.factor == "Checked")

data_clean$zone_count_c

data_clean$zone_category_c = cut(data_clean$zone_count_c, breaks = c(0, 1.5, 4.5, 6), 
                                labels = c("< 2", "2-4", "> 4"), include.lowest = TRUE, right = TRUE)


# A histogram to visualise this better
histogram(data_clean$zone_category_c)

# Proportions of records with any involvement in each of the three general zones (as a proportion of all participants):

# Upper zone (RUZ or LUZ)
upper_zone_proportion = mean(data_clean$consol_zones___1.factor == "Checked" | 
                              data_clean$consol_zones___4.factor == "Checked")

# Mid zone (RMZ or LMZ)
mid_zone_proportion = mean(data_clean$consol_zones___2.factor == "Checked" | 
                            data_clean$consol_zones___5.factor == "Checked")

# Lower zone (RLZ or LLZ)
lower_zone_proportion = mean(data_clean$consol_zones___3.factor == "Checked" | 
                              data_clean$consol_zones___6.factor == "Checked")

# Print the proportions
cat("Proportion of records with Upper zone involvement:", upper_zone_proportion, "\n")
cat("Proportion of records with Mid zone involvement:", mid_zone_proportion, "\n")
cat("Proportion of records with Lower zone involvement:", lower_zone_proportion, "\n")
```


```{r}

# Look at variable
data_clean$mech_vent.factor

```


```{r}

# The mech vent factor
data_clean$mech_vent.factor2 = data_clean$mech_vent.factor

data_clean$mech_vent.factor2[which(data_clean$mech_vent.factor == "Unknown/not reported")] = "No"

data_clean$mech_vent.factor2


```
```{r}

# The ICU admit factor creation
data_clean$icu_admit.factor2 = data_clean$icu_admit.factor

data_clean$icu_admit.factor2[which(is.na(data_clean$icu_admit.factor2))] = "No"

data_clean$icu_admit.factor2


```


```{r}

data_pcp = data_clean %>% filter(micro_result.factor == "Positive")
data_npcp = data_clean %>% filter(micro_result.factor != "Positive")

```

```{r}

data_clean$time_hivdx.factor

```



# Table value generators

```{r}
# An example for a categorical variable
# remove .factor for non-drop down variables 

library(rlang)

data_clean$time_hivdx.factor

variable = "time_hivdx.factor" # set your categorical variable here

# Use the describe function to see which categories are being included, then add the ones you don't want below, but don't remove
# the default.

filter_values = c("default", "Unknown/not recorded") # List the filtered out values here, separated by a comma. If you don't want to 
# filter, remove the category except for default.

data_cleaned = data_clean %>% filter(! (!!sym(variable)) %in% filter_values)

#cbind(data_clean$parenchy_yn___1.factor, data_cleaned$parenchy_yn___1.factor)

data_pcp = data_cleaned %>% filter(micro_result.factor == "Positive")
data_npcp = data_cleaned %>% filter(micro_result.factor != "Positive")

# Convert to unordered factor
data_cleaned[[variable]] = as.factor(data_cleaned[[variable]])

# The code below relevels factor variables - skip it entirely if you don't need it
data_cleaned[[variable]] = relevel(data_cleaned[[variable]], ref = "Prior to admission") # Replace New_Ref with the level of choice

##
## Don't change the code below, simply set the variable and run the cell
##

print("PCP Group:")
describe(data_pcp[[variable]]) # PCP group

print("Non-PCP Group")
describe(data_npcp[[variable]]) # Non-PCP group

# logistic regression
selected = data_cleaned[[variable]]
fit = glm(response ~ selected, data = data_cleaned, family = "binomial")

# fit = glm(response ~ selected + age + chd, data = data_cleaned, family = "binomial")

# extract the coefficients
coefficients = coef(fit)

# compute the Odds Ratios
OR = exp(coefficients)

print("OR")
OR

# compute the 95% confidence intervals
CI = exp(confint(fit))

# 95% conf ints
print("Confidence Interval")
CI

# p-value
sum = summary(fit) # P-value is under the Pr(>|z|) column (not the Intercept col)
sum$coefficients

```


```{r}

summary(data_pcp$severity.score)

```


You can do this for each of the variables you would like to assess. In your case, you would pay attention to the OR and confidence interval values associated with sex.factorFemale, which are relative to the male group in this case. Ie relative to the males, the odds of having PCP as a female are about 1.5 times higher than for males (unadjusted). The odds are about 52.42\% higher for females than for the males group.





```{r}
# An example for age (continuous variable)

data_clean$severity.score

data_clean$alb_new

variable = "severity.score" # set your continuous variable here

data_clean[[variable]] = as.numeric(data_clean[[variable]])

filter_values = c(Inf, -Inf) # List the filtered out values here if you need them. If you want values less than and greater than specific values, respectively.

data_cleaned = data_clean



#data_cleaned[[variable]] = data_clean %>% dplyr::filter(data_cleaned[[variable]] < Inf & 
#                                       data_cleaned[[variable]] > -Inf)

# data_cleaned[[variable]]

```


```{r}
#data_cleaned$severity.score
```


```{r}
data_pcp = data_cleaned %>% filter(micro_result.factor == "Positive")
data_npcp = data_cleaned %>% filter(micro_result.factor != "Positive")

##
## Don't change the code below, simply set the variable and run the cell
##
```



```{r}
print("PCP Group:")
pcp = describe(data_pcp[[variable]]) # PCP group
pcp
print(summary(data_pcp[[variable]]))

print("Non-PCP Group")
npcp = describe(data_npcp[[variable]]) # Non-PCP group
npcp
print(summary(data_npcp[[variable]]))

# Median difference
print("Median difference")
as.numeric(median(data_pcp[[variable]])) - as.numeric(median(data_npcp[[variable]]))

# Mann-Whitney U test
test_result = wilcox.test(data_pcp[[variable]], data_npcp[[variable]], exact = FALSE)

# Bootstrap for 95% CI of the median difference
set.seed(123)  # for reproducibility
n_iterations <- 10000
bootstrap_median_diffs <- numeric(n_iterations)

for(i in 1:n_iterations) {
  bootstrap_sample_pcp <- sample(data_pcp[[variable]], size = length(data_pcp[[variable]]), replace = TRUE)
  bootstrap_sample_npcp <- sample(data_npcp[[variable]], size = length(data_npcp[[variable]]), replace = TRUE)
  
  bootstrap_median_diffs[i] <- median(bootstrap_sample_pcp) - median(bootstrap_sample_npcp)
}

# Compute the 95% confidence interval
ci_lower <- quantile(bootstrap_median_diffs, 0.025, na.rm = T)  # 2.5 percentile
ci_upper <- quantile(bootstrap_median_diffs, 0.975, na.rm = T)  # 97.5 percentile

print(paste("95% Confidence Interval of the Median Difference: [", ci_lower, ", ", ci_upper, "]"))

# print the p-value from the bootstrap resampling test
print(paste("p-value: ", test_result$p.value))

```


```{r}
##
## Adjusted version of the OR
##

variable = "severity.score"

##
## Don't change the code below, simply set the variable and run the cell
##

# logistic regression
selected = data_clean[[variable]]
#fit = glm(response ~ selected, data = data_clean, family = "binomial")
fit = glm(response ~ selected + cld.factor + codiag, data = data_clean, family = "binomial")

# extract the coefficients
coefficients = coef(fit)
coefficients

# compute the Odds Ratios
OR = exp(coefficients)

print("OR")
OR

# compute the 95% confidence intervals
CI = exp(confint(fit))

# 95% conf ints
print("Confidence Interval")
CI

# p-value
sum = summary(fit) # P-value is under the Pr(>|z|) column (not the Intercept col)
sum$coefficients


```

This code performs bootstrap sampling 1000 times (you can increase this number for more precision) and then calculates the median difference for each bootstrap sample. The 2.5 and 97.5 percentiles of these bootstrap median differences are then used to form the 95\% confidence interval of the median difference.

Again, please note that the calculation of the confidence interval is under the assumption that the median difference follows the empirical distribution of the bootstrap samples, which might not be the case in reality.


# PCP vs Severe PCP for Table 5

```{r}

# Get rid of the NA values
data_clean$icu_admit.factor = ifelse(is.na(data_clean$icu_admit.factor), "No", data_clean$icu_admit.factor)
data_clean$ref_icu.factor = ifelse(is.na(data_clean$ref_icu.factor), "No", data_clean$ref_icu.factor)

# Create a severe PCP variable (composite) - This is just the initial variable - It will be adjusted in the code below - also do not know where the or | and and & should go below (unclear in the doc)
data_clean$severe_pcp = ifelse((data_clean$icu_admit.factor == "Yes" | data_clean$ref_icu.factor == "Yes" | 
                                data_clean$comb_pf <= 100 | data_clean$inhosp_death.factor == "Yes"), "Yes", "No")


describe(data_clean$severe_pcp)

data_pcp = data_clean %>% filter(micro_result.factor == "Positive" & severe_pcp == "No")
data_spcp = data_clean %>% filter(micro_result.factor == "Positive" & severe_pcp == "Yes")

```



```{r}
# An example for a categorical variable
# remove .factor for non-drop down variables 

library(rlang)

variable = "mech_vent.factor" # set your categorical variable here

# Use the describe function to see which categories are being included, then add the ones you don't want below, but don't remove
# the default.

filter_values = c("default", "Unknown/not recorded") # List the filtered out values here, separated by a comma. If you don't want to 
# filter, remove the category except for default.

data_cleaned = data_clean %>% filter(! (!!sym(variable)) %in% filter_values)

data_pcp = data_cleaned %>% filter(micro_result.factor == "Positive" & severe_pcp == "No")
data_spcp = data_cleaned %>% filter(micro_result.factor == "Positive" & severe_pcp == "Yes")

##
## Don't change the code below, simply set the variable and run the cell
##

print("PCP Group:")
describe(data_pcp[[variable]]) # PCP group

print("Severe-PCP Group")
describe(data_spcp[[variable]]) # Severe-PCP group

# logistic regression
selected = data_clean[[variable]]
fit = glm(response ~ selected, data = data_cleaned, family = "binomial")

# extract the coefficients
coefficients = coef(fit)

# compute the Odds Ratios
OR = exp(coefficients)

print("OR")
OR

# compute the 95% confidence intervals
CI = exp(confint(fit))

# 95% conf ints
print("Confidence Interval")
CI

# p-value
sum = summary(fit) # P-value is under the Pr(>|z|) column (not the Intercept col)
sum$coefficients

```


```{r}
# An example for age (continuous variable)

variable = "alb_new" # set your continuous variable here


filter_values = c(Inf, -Inf) # List the filtered out values here if you need them. If you want values less than and greater than specific values, respectively.

data_cleaned = data_clean %>% filter(variable < filter_values[1] & 
                                       variable > filter_values[2])

data_pcp = data_cleaned %>% filter(micro_result.factor == "Positive" & severe_pcp == "No")
data_spcp = data_cleaned %>% filter(micro_result.factor == "Positive" & severe_pcp == "Yes")

##
## Don't change the code below, simply set the variable and run the cell
##


print("PCP Group:")
pcp = describe(data_pcp[[variable]]) # PCP group
pcp

print("Severe-PCP Group")
spcp = describe(data_spcp[[variable]]) # Severe-PCP group
spcp

# Median difference
print("Median difference")
as.numeric(pcp$counts[10]) - as.numeric(spcp$counts[10])

# Mann-Whitney U test
test_result = wilcox.test(data_pcp[[variable]], data_spcp[[variable]], exact = FALSE)

# Bootstrap for 95% CI of the median difference
set.seed(123)  # for reproducibility
n_iterations <- 10000
bootstrap_median_diffs <- numeric(n_iterations)

for(i in 1:n_iterations) {
  bootstrap_sample_pcp <- sample(data_pcp[[variable]], size = length(data_pcp[[variable]]), replace = TRUE)
  bootstrap_sample_spcp <- sample(data_spcp[[variable]], size = length(data_spcp[[variable]]), replace = TRUE)
  
  bootstrap_median_diffs[i] <- median(bootstrap_sample_pcp) - median(bootstrap_sample_spcp)
}

# Compute the 95% confidence interval
ci_lower <- quantile(bootstrap_median_diffs, 0.025, na.rm = T)  # 2.5 percentile
ci_upper <- quantile(bootstrap_median_diffs, 0.975, na.rm = T)  # 97.5 percentile

print(paste("95% Confidence Interval of the Median Difference: [", ci_lower, ", ", ci_upper, "]"))

# print the p-value from the bootstrap resampling test
print(paste("p-value: ", test_result$p.value))

```

# Venn Diagram


```{r}

library(VennDiagram)

# TODO: Add nicer names for the files. Add the file save code too.

dataset = data_npcp # Set the dataset here (either data_pcp or data_npcp)

# Create lists for each condition
input = list(TB = dataset$record_id[which(dataset$tb == 1)], 
                  BacterialPneumonia = dataset$record_id[which(dataset$bacterialPneumonia == 1)], 
                  ViralPneumonia = dataset$record_id[which(dataset$viralPneumonia == 1)],
                  PulmonaryOedema = dataset$record_id[which(dataset$pulmonaryOedema == 1)],
                  Other = dataset$record_id[which(dataset$other == 1)])

#input

# Create the Venn diagram
venn.plot <- venn.diagram(x = input, filename = NULL, fill = c("skyblue", "pink1", "mediumorchid", "orange", "green"), alpha = 0.50,
                          label.col = "black", cex = 1.5, fontfamily = "serif", fontface = "bold",
                          cat.default.pos = "text", cat.default.cex = 1.5, cat.default.fontfamily = "serif", 
                          cat.default.fontface = "bold", cat.col = c("skyblue", "pink1", "mediumorchid", "orange", "green"))

?venn.diagram()

grid.draw(venn.plot)

```



These diagrams can actually be saved to a file externally if you need them to be high quality. Play around with colours etc if you would like.



